NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2019-02-06 Manuscript/manuscriptPS.tex
Stylistics tuning of introduction.

\section{Conclusions}
Lipids with PS headgroups and their interactions with ions
play an important role in lipid mediated signaling
processes \cite{leventis10,yeung08}. Recent studies using molecular
dynamics simulations to interpret the various spectroscopic data
give contradictory results for the calcium binding details to PS
headgroups \cite{melcrova16,valentine18,hallock18}.  
Here, we used the headgroup C-H bond order parameters and open collaboration method to evaluate the quality
of headgroup structure and ion binding affinity to PS lipids
in available MD simulation force fields,
as previously done for PC lipids \cite{botan15,catte16}.
The main advantage of this approach is the direct connection
between accurately measured experimental order parameters and simulations
which reduces the ambiguity in the interpretation of experiments.

First, we complemented the available experimental data of PS
lipid headgroup order parameters \cite{browning80,roux90} by measuring the signs of the order parameters.
None of the available force fields tested using the NMRlipids open
collaboration was accurate enough to reproduce the PS
headgroup order parameters with the experimental accuracy, but
the best models suggested a characteristic rigid conformation for the
carboxyl group in the serine headgroup. Comparison to the previously
measured headgroup order parameters from POPC:POPS (5:1) bilayers with different ion
concentrations \cite{roux90} showed that the tested MD simulation force fields
overestimate the cation binding affinity to the negatively charged bilayers
containing PS lipids with two exceptions. The apparently most realisitic monovalent ion binding
affinity to PS containing lipid bilayers was observed in the MacRog
simulations with potassium counterions and the CHARMM36 force field with recently introduced
NBfix correction for calcium \cite{kim16} underestimated the calcium binding affinity.
However, the experimentally measured trends of the PS headgroup order parameter response
to the bound calcium and to the dilution of bilayer with zwitterionic PC lipids were not
qualitatively reproduced in any of the tested force fields, indicating that improvements in
the MD simulation force fields are necessary to study interactions between PS lipids and other biomolecules.
This is different to the previous results with PC lipids,
where the experimentally measured headgroup order parameter responses to the bound charge
were qualitative reproduced even thought the headgroup structures without ions were not
correct and the cation binding affinities were overestimated \cite{catte16}.

Our results pave the way for the development of better MD simulations force fields
for PS lipids. Using the headgroup order parameters, we were able to evaluate the
quality of various conformational ensembles in different force fields.
This can guide the development of force fields that would correctly reproduce the
conformations sampled by PS headgroups. The experimental dataset of headgroup order
parameters from POPC:POPS (5:1) mixture with different cation concentrations can
be used to improve cation binding details in the force fields, as recently demonstrated
for POPC using the electronic continuum correction \cite{melcr18}.
Similar study for POPS is progressed separately \cite{ECCpops}.

>>>>

\section{Conclusions}
Lipids with PS headgroups, and their interactions with ions,
play an important role in lipid-mediated signaling
processes \cite{leventis10,yeung08}. Recent studies using MD
simulations to interpret the various spectroscopic data
give contradictory results for the calcium binding details to PS
headgroups \cite{melcrova16,valentine18,hallock18}.  
Here, as was previously done for PC lipids \cite{botan15,catte16},
we used the headgroup C--H bond order parameters and the open collaboration approach to evaluate the quality
of the headgroup structure and ion binding affinity to PS lipids
in available MD force fields.
The main advantage of this approach is the direct connection
between the accurately measured experimental order parameters and the simulations,
which reduces the ambiguity in the interpretation of experiments.

First, we complemented the available experimental data of PS
lipid headgroup order parameters \cite{browning80,roux90} by measuring the signs of the order parameters.
Comparison to these data revealed that none of the available force fields
%tested using the NMRlipids open collaboration
was accurate enough to reproduce the PS
headgroup order parameters within the experimental accuracy. However,
the best models for the serine headgroup suggested a characteristic rigid conformation for its
carboxyl group. Comparison to the previously
measured headgroup order parameters from POPC:POPS (5:1) bilayers with different ion
concentrations \cite{roux90} then showed that the tested MD force fields
overestimate the cation binding affinity to the negatively charged bilayers
containing PS lipids with two exceptions. 1) The apparently most realistic monovalent ion binding
affinity to PS-containing lipid bilayers was observed in the MacRog
simulations with potassium counterions. 2) The CHARMM36 force field with the recently introduced
NBfix correction for calcium \cite{kim16} underestimated the calcium binding affinity.
The experimentally measured trends of the PS headgroup order parameter response
to the bound calcium, and to the dilution of bilayer with zwitterionic PC lipids, were not
qualitatively reproduced in any of the tested force fields, which indicates that improvements in
the MD force fields are necessary to study interactions between PS lipids and other biomolecules.
This is different to the previous results with PC lipids,
where the experimentally measured headgroup order parameter responses to the bound charge
were qualitative reproduced even though the headgroup structures themselves were
incorrect and the cation binding affinities overestimated \cite{catte16}.

Our results pave the way for the development of better MD force fields
for PS lipids. Using the headgroup order parameters, we were able to evaluate the
quality of various conformational ensembles in different force fields.
This can guide the development of force fields that would correctly reproduce the
conformations sampled by PS headgroups. The experimental dataset of headgroup order
parameters from POPC:POPS (5:1) mixture with different cation concentrations can
be used to improve cation binding details in the force fields, as recently demonstrated
for POPC using the electronic continuum correction \cite{melcr18}.
Similar study for POPS is being progressed separately \cite{ECCpops}.
